Boosted Mirror Descent, Frank-Wolfe, and Herding

### 2013-09-13 icml/paper.tex
small typo fixes

%%% ALGORITHM 1 ANALYSIS %%%
% Convergence proposition
\begin{proposition}[Convergence of Algorithm 1]
\label{prop:method-1}
Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
Then we have
\begin{equation}
\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}).
\end{equation}
\end{proposition}
\begin{proof}
Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
$z_t = \theta_t$, then we get the inequality
\begin{align}
\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}). \nonumber
\end{align}
Combining these together, we get the string of inequalities
\begin{align*}
L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_f(u_t \| u_{t+1}) \\
 &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_f(u_t \| u_{t+1}),
\end{align*}
as was to be shown.
\end{proof}

% Convergence rate theorem
\begin{theorem}
\label{thm:method-1}
Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\begin{proof}
By Lemma~\ref{lem:convexity}, we have 
\begin{align*}
D_f(x_t \| x_{t+1}) &= D_{f^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
 &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
 &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
 &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
 &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
 &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
\end{align*}
It follows that 
\begin{align*}
\frac{1}{A_T} \sum_{t=1}^T A_tD_f(x_t \| x_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
\end{align*}
\end{proof}

@@

\NA{JHH: Need to specify the spaces we are working with here.}
Suppose we are given a family of features $\phi$ together 
with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
would like to construct a distribution that approximately 
matches this distribution, i.e.~one for which 
$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
to do this is by using boosted mirror descent to minimize the 
maximum mean discrepancy relative to $\phi$; 
in other words, to minimize
\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
If we wish to write this more similarly to the original problem 
formulation, we can write
\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
the updates are the same as the standard herding updates. Let $r = \sup_{x} \|\phi(x)\|_2$; 
then using the Algorithm 2 convergence bound gives us
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{8r^2(\log(T)+1)}{T} \\
 &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{8R^2(\log(T)+1)}{T} \\
 &= \frac{8R^2(\log(T)+1)}{T}.
\end{align*}
This is a slightly weaker version of the typical herding bound on MMD.

>>>>

%%% ALGORITHM 1 ANALYSIS %%%
% Convergence proposition
\begin{proposition}[Convergence of Algorithm 1]
\label{prop:method-1}
Consider the updates $\theta_t \in \arg\max_{\theta} \langle \theta, u_t \rangle - R(\theta)$ 
and $u_{t+1} \in \arg\min_u h(u) + \langle \hat{\theta}_s, u \rangle$. 
Then we have
\begin{equation}
\sup_{\theta} L(\hat{u}_T, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}).
\end{equation}
\end{proposition}
\begin{proof}
Note that $L(u_t, \theta_t) = \max_{\theta} L(u_t, \theta)$ by construction. 
Also note that, if we invoke Lemma~\ref{lem:bregman} with $f = h$ and 
$z_t = \theta_t$ then we get the inequality
\begin{align}
\frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) 
&\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) \\
&\quad + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(x_t \| x_{t+1}). \nonumber
\end{align}
Combining these together, we get the string of inequalities
\begin{align*}
L(\hat{u}_T, \theta) &= L\left(\frac{1}{A_T} \sum_{t=1}^T \alpha_tu_t, \theta\right) \\
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta) \\
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u_t, \theta_t) \\
 &\leq \frac{1}{A_T} \sum_{t=1}^T \alpha_t L(u^*, \theta_t) + \frac{1}{A_T} \sum_{t=1}^T A_t D_h(u_t \| u_{t+1}) \\
 &\leq \sup_{\theta} L(u^*, \theta) + \frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}),
\end{align*}
as was to be shown.
\end{proof}

% Convergence rate theorem
\begin{theorem}
\label{thm:method-1}
Suppose that $h$ is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{\theta} \|\theta\|_{*}$. Then 
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\begin{proof}
By Lemma~\ref{lem:convexity}, we have 
\begin{align*}
D_h(u_t \| u_{t+1}) &= D_{h^*}(\hat{\theta}_{t+1} \| \hat{\theta}_{t}) \\
 &\leq \frac{1}{2}\|\hat\theta_{t+1}-\hat\theta_{t}\|_{*}^2 \\
 &= \frac{1}{2}\|\frac{\sum_{s \leq t}\alpha_s\theta_s}{\sum_{s \leq t} \alpha_s} - \frac{\sum_{s \leq t+1}\alpha_s\theta_s}{\sum_{s \leq t+1} \alpha_s}\|_{*}^2 \\
 &= \frac{1}{2}\|\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\theta_s - \frac{\alpha_{t+1}}{A_{t+1}} \theta_{t+1}\|_{*}^2 \\
 &\leq \frac{1}{2} \left(\frac{\alpha_{t+1}}{A_tA_{t+1}} \sum_{s \leq t} \alpha_s\|\theta_s\|_{*}^2 + \frac{\alpha_{t+1}}{A_{t+1}} \|\theta_{t+1}\|_{*}^2\right)^2 \\
 &\leq \frac{2r^2\alpha_{t+1}^2}{A_{t+1}^2}.
\end{align*}
It follows that 
\begin{align*}
\frac{1}{A_T} \sum_{t=1}^T A_tD_h(u_t \| u_{t+1}) &\leq \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}.
\end{align*}
\end{proof}

@@

\NA{JHH: Need to specify the spaces we are working with here.}
Suppose we are given a family of features $\phi$ together 
with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
would like to construct a distribution that approximately 
matches this distribution, i.e.~one for which 
$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
to do this is by using boosted mirror descent to minimize the 
maximum mean discrepancy relative to $\phi$; 
in other words, to minimize
\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
If we wish to write this more similarly to the original problem 
formulation, we can write
\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
the updates are the same as the standard herding updates. Let $r = \sup_{x} \|\phi(x)\|_2$; 
then using the Algorithm 2 convergence bound gives us
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T} \\
 &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{8r^2}{T} \\
 &= \frac{8r^2}{T}.
\end{align*}
This is the same rate as the typical herding bound on MMD.
